### 2017-10-28 Risk analysis/Text/Layout/Layout/chapter2_methods.tex

@@

        	\item \textbf{Seal} formations that are able to halt hydrocarbons on their migration pathway as top, lateral and possibly bottom seals \citep{dolson2016basics, sorkhabi2005place}. Such sealing is generally defined by a reduction of pore space geometry relative to migration energy. Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats.These are characterized by a high capillary entry pressure able to resist the buoyancy pressure of hydrocarbons. Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts \citep{dolson2016basics, sorkhabi2005place, dolson2016quantifying}. An additional factor to sealing is posed by the presence of faults. Fault zones may possess significantly higher or lower permeabilities then surrounding rock formations which means that they may serve as fluid flow conduits or migration barriers respectively \citep{van2003lateral, sorkhabi2005place}. The possible influence of faults on structural hydrocarbon traps is elaborated in further detail below. For the model in this work, the reservoir cap rock is assumed to be a shale.
        	\item Suitable \textbf{reservoir formations}, in which these hydrocarbons volumes can be stored. For this, a suitable percentage of void space in a rock formation, i.e. a sufficiently high porosity, is required. A second significant factor is the connectivity of the porosity through pore throats. The larger the pore throats, the easier it is for fluids to migrate in a rock. Typical reservoir rocks are sandstones and carbonates, but even normally tight lithologies might serve as reservoirs, if they are fractured \citep{dolson2016basics}. In the models of this work, reservoir formations are assumed to be conventional sandstones.

@@

        Given the migration and a combination of appropriate reservoir-seal sequence geometries, many different types of traps, in which hydrocarbons can accumulate, are possible. A common trap type is a structural four-way closure, i.e. as reservoir-seal system that was folded into closed anticlines, often resembling a dome and is thus predominantly defined by the structural closure relief and the cap rock sealing capacity. Such a trap type can be considered to be closed, as the reservoir surface contours are closed in a circle or oval around a relief high. Such closure can also be given through any other crossing of a structural contour with a seal, including faults. A point at which a contour fails to close is referred to as "spill point" \citep{dolson2016basics}.

@@

        	\includegraphics[width=0.95\textwidth]{Figures/trap_closure.png}
        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief countour) and cross section perspective (modified from \citet{dolson2016basics}.)}

@@

        The influence of faults and fault sealing potential on structural traps has been outlined by several researches such as \citet{sorkhabi2005place} and \citet{van2003lateral}.

@@

				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA.png}

@@

				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB.png}

@@

			\caption{Bla}\label{fig:fault_trap_spills}

@@

        In this work, reservoir systems and traps resulting from modeling are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e. Bayesian decision theory is applicable. In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.

>>>>

        \subsection{Bayesian hierarchical networks}\label{sec:hierarchical_bayes}

@@

        	\item \textbf{Seal} formations that are able to halt hydrocarbons on their migration pathway as top, lateral and possibly bottom seals \citep{dolson2016basics, sorkhabi2005place}. Such sealing is generally defined by a reduction of pore space geometry relative to migration energy. Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats. These are characterized by a high capillary entry pressure able to resist the buoyancy pressure of hydrocarbons. This capillary pressure is also the main controlling factor for seal capacity. Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts \citep{dolson2016basics, sorkhabi2005place, dolson2016quantifying}. An additional factor to sealing is posed by the presence of faults. Fault zones may possess significantly higher or lower permeabilities then surrounding rock formations which means that they may serve as fluid flow conduits or migration barriers respectively \citep{van2003lateral, sorkhabi2005place}. The possible influence of faults on structural hydrocarbon traps is elaborated in further detail below. For the model in this work, the reservoir cap rock is assumed to be a shale.
        	\item Suitable \textbf{reservoir formations}, in which these hydrocarbons volumes can be stored. For this, a suitable percentage of void space in a rock formation, i.e. a sufficiently high porosity, is required. Further significant factors are the connectivity of the porosity through pore throats, as well as pore geometry and resulting permeability \citep{dolson2016basics, sorkhabi2005place}. The larger the pore throats, the easier it is for fluids to migrate in a rock. Typical reservoir rocks are sandstones and carbonates, but even normally tight lithologies might serve as reservoirs, if they are fractured \citep{dolson2016basics}. In the models of this work, reservoir formations are assumed to be conventional sandstones.
        	\item A \textbf{trap} resulting from an appropriate reservoir-seal sequence and a three-dimensional geometric closure. Regarding a top view on a structural relief, closure is generally defined by any intersection of a structural contour and the seal, that is closed on both sides \citep{dolson2016basics, sorkhabi2005place}. A point at which a contour fails to close is referred to as "spill point". The maximum column depth to which a trap can be filled, and ultimately its maximum volume, is defined by this spill point. A trap that contains hydrocarbons down to the spill point is thus referred to as "filled to spill". traps can also be "filled to seal capacity" when the seal capacity is exceeded by the buoyant pressure of a fill column of greater height. Lower accumulations can also result from insufficient migration, i.e. low charge volumes of oil or gas in the first place. Such a trap is referred to as "charge limited" \citep{dolson2016basics}. Many different types of traps, in which hydrocarbons can accumulate, are possible. A common trap type is a structural four-way closure, i.e. as reservoir-seal system that was folded into closed anticlines, often resembling a dome and is thus predominantly defined by the structural closure relief and the cap rock sealing capacity. Such a trap type can be considered to be closed, as the reservoir surface contours are closed in a circle or oval around a relief high. Closure can also be given through the intersection of a fault with the seal \citep{dolson2016basics}. Two common examples are illustrated in Figure \ref{fig:trap_types}.

@@

@@

        	\includegraphics[width=0.95\textwidth]{Figures/trap_closure.PNG}
        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.)}\label{fig:trap_types}

@@

        %The influence of faults and fault sealing potential on structural traps has been outlined by several researches such as \citet{sorkhabi2005place} and \citet{van2003lateral}.
        The presence of a fault in the reservoir-seal sequence can have different impacts on trap closure and the volume of a hydrocarbon accumulation. This is mainly defined by the transport properties of a fault, i.e. in which way it behaves as a seal or flow conduit respectively. Cases might range from complete fault sealing to complete leakage along a fault zone. More complex in-between cases are possible, such as faults that leak across, but are sealed along their plane. This ultimately defines the depth to which a trap may be filled and whether this point is controlled by a point of leakage on the fault plane or a spill point related to the remainder of the 3D structural closure. Another decisive aspects to consider, it the nature of juxtapositions across the fault, especially regarding the possibility of cross-fault leakage \citep{van2003lateral}. This is summarized in Figure \ref{fig:fault_trap_spills}.\\

@@

				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA.PNG}

@@

				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB.PNG}

@@

			\caption{In (a), a structural trap resulting from the combination of 4-way anticlinal closure and a simple normal fault is illustrated. The two reservoir layers in both sides are juxtaposed, so that the hydrocarbon accumulation potential is controlled by three points of possible leak pathways: (1) the anticlinal spill point, (2) leakage upwards along the fault and (3) leakage across the fault enabled by the juxtaposition. Possibilities (2) and (3) depend on the transport properties of the fault zone, i.e. potential fault sealing. Given leakage along the fault (2), that size volume would be reduced to a small relatively small volume solely defined by the 4-way closure down to maximum contact of the reservoir-seal boundary with the fault plane. In case (b), the reservoirs are not juxtaposed and thus laterally sealed. Fault-related leakage can thus only occur along the plane. Leakage along and across the fault is enabled by the juxtaposition in (c) (modified from \citet{van2003lateral}).}\label{fig:fault_trap_spills}

@@

		The transport properties of a fault zone are primarily defined by the internal structure and composition of its fault gouge. Fault gouge properties and ultimately the variation in permeability in the fault zone is primarily dependent on clay content. Clay smearing has thus been a central topic of research surrounding fault sealing (see \citep{yielding1997quantitative, van2003lateral, van2005processes, schmatz2010clay}. It generally is defined as all processes that lead to the incorporation of clay from the wall rock (e.g. a shale as seal) into the fault zone \citet{van2003lateral, vrolijk2016clay}. Clay smearing can be described using different equations, such as the Shale Smear Factor and the Shale Gouge Ratio \citep{yielding1997quantitative, vrolijk2016clay}. These are primarily based on the relation of displaced shale to fault throw (IF I STILL GET TO USE THESE, ELABORATE!). A critical review of the respective research published over the last four decades was recently presented by \citet{vrolijk2016clay}.\\		
        The 3D structural model in this work (see Section \ref{sec:3D_design}) is designed in a way that a potential conventional hydrocarbon trap is included in the form of anticlinal folding combined with normal faulting of a reservoir-seal sequence. Results from modeling are to be evaluated in a way that respective estimations on the basis of loss functions can be conducted, i.e. Bayesian decision theory is applicable. In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}. Thus, several of the factors named above are to be considered to attain a calculable volume of the structural trap which can further be economically interpreted using the approaches presented below.
